Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue

### 2021-01-15 reports/rmarkdown/03_integrated_datasets.Rmd
adds fig2 and figS1-2

#Graph construction  
k <- 15 # k for KNN
P <- 0.1# prunning threshold of SNN

@@

#Compute SNN graph from KNN
jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))

jSNN[1:10,1:10]

# k=10
# P=0.1
snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
snnj <- matrix(0, nrow(snn$id), nrow(snn$id))

# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }

for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }

snnm[ snnj <= ceiling(k*P) ] <- 0

#Prune SNN graph based on the Jaccard index
pSNN <- jSNN
pSNN <- pSNN / max(pSNN)

@@

#Graph construction 
k_nearest_neighbours <- 10 # k for KNN
snn_threshold <- 0.1 # prunning threshold of SNN

>>>>

#Graph construction  
k <- 15 # k for KNN
P <- 0.2# prunning threshold of SNN

@@

#Compute SNN graph from KNN
# jSNN <- t( t( (nn %*% nn ) / rowSums(nn)) / colSums(nn))
# 
# jSNN[1:10,1:10]
# 
# # k=10
# # P=0.1
# snn <- dbscan::sNN(x = adj, k = k, jp = T, approx = TRUE)
# snnm <- matrix(0, nrow(snn$id), nrow(snn$id))
# snnj <- matrix(0, nrow(snn$id), nrow(snn$id))
# 
# # for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$dist[idx,] }
# for(idx in 1:nrow(snn$id) ){  snnm[idx,snn$id[idx,]] <- snn$shared[idx,]/k }
# 
# for(idx in 1:nrow(snn$id) ){  snnj[idx,snn$id[idx,]] <- snn$shared[idx,] }
# 
# snnm[ snnj <= ceiling(k*P) ] <- 0

#Prune SNN graph based on the Jaccard index
# pSNN <- jSNN
# pSNN <- pSNN / max(pSNN)

@@

#Graph construction 
k_nearest_neighbours <- 5 # k for KNN
snn_threshold <- 0.1 # prunning threshold of SNN

### 2021-01-15 reports/rmarkdown/manuscript/manuscript_template.Rmd
adds fig2 and figS1-2

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(ggplot2)
library(igraph)
library(dplyr)
library(stringr)
library(rafalib)
# remotes::install_github("czarnewski/niceRplots")
library(niceRplots)

@@

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
par(mar=c(4,5,1,12))
temp <- t(t(2^x-1)/colSums(2^x-1))
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"
taxa <- taxa[rownames(temp),]
temp <- rowsum(temp , taxa$Class)
temp[is.nan(temp)] <- 0
o <- order( rowSums(temp) , decreasing = T)
sample_o <- order( patient_groups )
temp <- temp[o,sample_o]*100
temp <- temp[-grep("NA",rownames(temp)),]
temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

@@

@@

@@

@@

>>>>

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(ggplot2)
library(igraph)
library(dplyr)
library(stringr)
library(rafalib)
library(scales)
# remotes::install_github("czarnewski/niceRplots")
library(niceRplots)

@@

#######################
# BACTERIAL ABUNDANCE #
#######################
par(mar=c(4,5,1,12))
temp <- t(t(2^x-1)/colSums(2^x-1))
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"
taxa <- taxa[rownames(temp),]
temp <- rowsum(temp , taxa$Class)
temp[is.nan(temp)] <- 0
o <- order( rowSums(temp) , decreasing = T)
sample_o <- order( patient_groups )
temp <- temp[o,sample_o]*100
temp <- temp[-grep("NA",rownames(temp)),]
temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

@@

```{r, echo = FALSE, eval = TRUE}
# Define Layout and set labels
layout(matrix(c(2 ,3,4,5,      6,7,20,20,          20,20,20,20,
                19,1,1,1,      1,8,20,20,          20,20,20,20,
                18,1,1,1,      1,9,21,21,         21,21,24,24,
                17,1,1,1,      1,10,22,22,         22,22,24,24,
                16,15,14,13,  12,11,23,23,         23,23,24,24 ) , ncol= 12,byrow = T),
       widths = 1,heights = c(1.2,1.2,1.2,1.2,1.2))
# layout.show(12)
figlabels <- letters

#########################
# BACTERIAL COMMUNITIES #
#########################
gB <- graph_from_adjacency_matrix(as.matrix(SNN_bacteria), mode = "undirected",diag = F,weighted = T)
set.seed(1)
lB <- layout_with_fr(gB,niter=3000,start.temp=30)

par(mar=c(0,0,0,0))
plot( gB , vertex.label.cex=0.000001 , vertex.color = pal[factor(bac_communities)] ,
      edge.width=  ( E(gB)$weight / max(E(gB)$weight)) ,
      vertex.size=10,main="", 
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(gB)$weight / max(E(gB)$weight) * 88 )+1 ] )
legend("topright",title.adj = 0,
       legend = levels(factor(bac_communities)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)

par(mar=c(1,1,1,1))
for(j in sort(unique(bac_communities)) ){
  empty_plot()
  text( 0,1,labels = paste0(rownames(SNN_bacteria)[ bac_communities == j ],collapse = "\n"),adj=c(0,1),cex=.5)
  title(main=paste0("community",j),cex.main=.7)
}
for(i in 1:(18-length(unique(bac_communities)))){empty_plot()}

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"

par(mar=c(1,7,2,3))
x <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),2]
y <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),3]
# plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
#        plot_labels = T, plot_weights = F,order_1_by = "total",
#        order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
#        pal=hue_pal()(length(unique(x))))

# plot_sankey( data.frame( y[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
#      plot_labels = T, plot_weights = F,order_1_by = "total",
#      order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
#      pal=hue_pal()(length(unique(y))))

plot_sankey( data.frame( x[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
     plot_labels = T, plot_weights = F,order_1_by = "total",
     order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
     pal=hue_pal()(length(unique(x))))

add_letter(figlabels[1]); figlabels <- figlabels[-1]

##############################################
# BACTERIAL COMMUNITY DIFFERENTIAL ABUNDANCE #
##############################################
par(mar=c(1,7,1,3))

module_means_per_dataset <- lapply(datasets_all_samples[-1],function(x){
  return( rowsum( x[rownames(SNN_bacteria),] , group = bac_communities ) / c(table(bac_communities)) )
})
module_means_per_dataset <- lapply(module_means_per_dataset,function(i){
  rownames(i) <- paste0("BC",rownames(i))
  return(i)
})

for( j in c("ASV_tissue_V3_normalized_batch_corrected",
        "ASV_CVL_V3_normalized_batch_corrected",
        "ASV_CVL_V2_normalized_NOT_batch_corrected")){
  temp_group <- as.character(patient_groups)
  temp_group[ is.na(temp_group) ] <- "NA"
  temp_group[ temp_group == "" ] <- "NA"
  temp_group <- factor(temp_group)

  barlist(data=module_means_per_dataset[[j]],
          genes=rownames(module_means_per_dataset[[j]]),col = pal,
          clustering=temp_group, main=paste0("\n", sub("_norm.*","",j)))
}
add_letter(figlabels[1]); figlabels <- figlabels[-1]

empty_plot()

@@

```{r, echo = FALSE, eval = TRUE}
# 
# figlabels <- letters
# temp_legend <- NULL
# 
# layout(matrix(c(1,4,
#                 2,4,
#                 3,4) , ncol= 2,byrow = T),widths = c(4,1),heights = c(1,1,1))
# 
# for(i in c("ASV_CVL_V3_normalized_batch_corrected",
#            "ASV_CVL_V2_normalized_NOT_batch_corrected",
#            "ASV_tissue_V3_normalized_batch_corrected")){
#   x <- datasets_all_samples[[i]]
#   
#   par(mar=c(3,5,1,1))
#   temp <- t(t(2^x-1)/colSums(2^x-1))
#   taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
#   taxa[is.na(taxa)] <- "NA"
#   taxa <- taxa[rownames(temp),]
#   temp <- rowsum(temp , taxa$Class)
#   temp[is.nan(temp)] <- 0
#   
#   if(is.null(temp_legend)){ 
#     temp_legend <- temp   #Use the same legend from the 1st dataset
#   }
#   
#   o <- order( rowSums(temp_legend) , decreasing = T)
#   sample_o <- order( patient_groups )
#   temp <- temp[o,sample_o]*100
#   temp <- temp[-grep("NA",rownames(temp)),]
#   temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )
#   
#   taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
#   barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
#            col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
#   # legend(x = par("usr")[c(2)],
#   #        y = par("usr")[c(4)],title.adj = 0,
#   #        legend = rownames(temp),xjust = 0,yjust = 1,
#   #        bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
#   #        pt.bg = taxa_pal)
#   
#   #add label
#   add_letter(figlabels[1]); figlabels <- figlabels[-1]
# }
# 
# 
# par(mar=c(4,0,1,0))
# empty_plot()
# legend(x = par("usr")[c(1)],
#          y = par("usr")[c(4)],title.adj = 0,
#          legend = rownames(temp),xjust = 0,yjust = 1,
#          bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
#          pt.bg = taxa_pal)

figlabels <- letters
temp_legend <- NULL

layout(matrix(c(1,7,
                2,7,
                3,7,
                4,7,
                5,7,
                6,7) , ncol= 2,byrow = T),widths = c(4,1),heights = c(.5,1,.5,1,.5,1))

for(i in c("ASV_CVL_V3_normalized_batch_corrected",
           "ASV_CVL_V2_normalized_NOT_batch_corrected",
           "ASV_tissue_V3_normalized_batch_corrected")){
  x <- datasets_all_samples[[i]]

  par(mar=c(1,5,1,1))
  temp <- t(t(2^x-1)/colSums(2^x-1))

  shann <- vegan::diversity(t(temp),index = c("shannon") )
  simp <- vegan::diversity(t(temp),index = c("simpson") )
  invsimp <- vegan::diversity(t(temp),index = c("invsimpson") )
  divers <- as.matrix(rbind(shann,simp,invsimp) )
  divers[is.infinite(divers)] <- 0
  rownames(divers) <- c("Shannon","simpson","Inv Simpson")
  barlist( data = divers,
         main = "",
         genes = c("Shannon","simpson","Inv Simpson"),
         clustering = patient_groups,
         draw_mean_lines=F,col = pal)

  par(mar=c(3,5,1,1))
  # taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
  # taxa[is.na(taxa)] <- "NA"
  # taxa <- taxa[rownames(temp),]
  # temp <- rowsum(temp , taxa$Class)
  temp[is.nan(temp)] <- 0

  if(is.null(temp_legend)){ 
    temp_legend <- temp   #Use the same legend from the 1st dataset
  }

  o <- order( rowSums(temp_legend) , decreasing = T)
  sample_o <- order( patient_groups )
  temp <- temp[o,sample_o]*100
  # temp <- temp[-grep("NA",rownames(temp)),]
  temp <- rbind(temp[1:20,], Other = colSums(temp[21:nrow(temp),]) )

  taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
  barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",main=sub("_nor.*","",i),font.main=1,
           col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] )
  # legend(x = par("usr")[c(2)],
  #        y = par("usr")[c(4)],title.adj = 0,
  #        legend = rownames(temp),xjust = 0,yjust = 1,
  #        bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
  #        pt.bg = taxa_pal)

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]
}

par(mar=c(4,0,1,0))
empty_plot()
legend(x = par("usr")[c(1)],
         y = par("usr")[c(4)],title.adj = 0,
         legend = rownames(temp),xjust = 0,yjust = 1,
         bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
         pt.bg = taxa_pal)
```

@@

```{r, echo = FALSE, eval = TRUE}

figlabels <- letters
temp_legend <- NULL

mypar(1,4,mar=c(12,1,1,0))

for(i in c("ASV_CVL_V3_normalized_batch_corrected",
           "ASV_CVL_V2_normalized_NOT_batch_corrected",
           "ASV_tissue_V3_normalized_batch_corrected")){

  continuous_parameters <- c("age","Count_CD4_v2","Count_CD4_v3","Plasma_S_Prog_nmol_L_v3","Plasma_S_Estradiol_pmol_L_v3","Conc_T4_v2","Conc_T3_v2","Conc_Progesteron_v2","Conc_Estradiol_v2","Conc_Cortisol_v2","Freq_CD4_v2","Freq_CD4_v3","Freq_Lympho_v2","Freq_Lympho_v3","visit_clients7days_v2")
  continuous_metadata <- metadata[,continuous_parameters]
  continuous_metadata <- cbind(continuous_metadata , t(module_means_per_dataset[[i]]) )

  metadata_cor <- cor(continuous_metadata,use = "pairwise.complete.obs")

  #TO ORDER THE HEATMAP
  o <- hclust( as.dist(1-metadata_cor),method = "complete")$order
  metadata_cor <- metadata_cor[o,o]

  image( t(metadata_cor [nrow(metadata_cor):1,]),axes=F,
         breaks = seq(-1,1,length.out = 90),main=sub("_nor.*","",i),font.main=1,
         col=colorRampPalette(c("navy","grey95","firebrick"))(89) )
  # text(  par("usr")[2]+.02 ,seq(1,0,length.out = nrow(metadata_cor)) , 
  #        labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
  text( seq(0,1,length.out = nrow(metadata_cor)) , par("usr")[3]-.02 , 
        labels = rownames(metadata_cor), srt = 90, adj = c(1,.5), xpd = TRUE, cex=1)

  #add label
  add_letter(figlabels[1]); figlabels <- figlabels[-1]
}

par(mar=c(12,0,1,0))
image( t(metadata_cor [nrow(metadata_cor):1,]),axes=F,
         breaks = seq(-1,1,length.out = 90),main="",
         col=colorRampPalette(c("white"))(89) )
text(  0 ,seq(1,0,length.out = nrow(metadata_cor)) , 
  labels = rownames(metadata_cor), srt = 0, adj = c(0,.5), xpd = TRUE, cex=1)
```

@@

```{r, echo = FALSE, eval = TRUE}
```

### 2021-01-15 reports/rmarkdown/manuscript/manuscript_template.tex
adds figure1 in manuscript
updates manuscript
adds fig2 and figS1-2

\title{Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A demonstration of Rmarkdown using Herman Bumpus' data}
\author{Author One\textsuperscript{1},
Author Two\textsuperscript{2},
Author Three\textsuperscript{1,2}}
\date{December 16, 2020}

@@

To be able to better undertand the differences in microbiome profile across all datasets collected, we performed a joint graph-based clustering analysis in order to identify co-regulated bacterial communities (see ``Methods'' section for details). A total of 11 bacterial communities were identified.

Noticebly, bacterial community 11 consisted only of Lactobacillus species (\emph{Anaerococcus hydrogenalis, Atopobium deltae, Bifidobacterium adolescentis, Bifidobacterium bifidum, Bifidobacterium longum/breve, Brevibacterium paucivorans, Brevibacterium ravenspurgense, Campylobacter hominis, Chlamydia trachomatis, Corynebacterium aurimucosum, Corynebacterium genitalium, Corynebacterium imitans, Corynebacterium jeikeium, Corynebacterium pseudogenitalium, Corynebacterium simulans/striatum, Corynebacterium sp/amycolatum, Corynebacterium spheniscorum, Corynebacterium sundsvallense/thomssenii, Dermabacter hominis, Dorea formicigenerans, Enterococcus villorum/hirae/faecium/faecalis/canintestini/canis/dispar/durans, Gemmiger formicilis, Lactobacillus coleohominis, Lactobacillus crispatus, Lactobacillus crispatus/acidophilus, Lactobacillus gasseri/johnsonii/taiwanensis, Lactobacillus iners, Lactobacillus jensenii, Lactobacillus mucosae, Lactobacillus reuteri, Lactobacillus reuteri/oris/frumenti/antri, Lactobacillus salivarius, Megasphaera elsdenii, other Bacilli, other Bifidobacteriaceae, other Clostridia, other Corynebacterium, other Ezakiella, other Gemella, other Gordonia, other Lactobacillales, other Lactobacillus, other Listeria, other Mobiluncus, other Mogibacterium, other Murdochiella, other Nesterenkonia, other Paraeggerthella, other Rhizobium, Peptoniphilus asaccharolyticus, Prevotella denticola, Staphylococcus saccharolyticus/capitis/hominis/caprae/lugdunensis/epidermidis/simiae, Streptococcus dysgalactiae, Streptococcus lutetiensis/equinus, Ureaplasma parvum, Brevundimonas mediterranea/intermedia/poindexterae, Corynebacterium appendicis/imitans, Corynebacterium freneyi/sp/amycolatum, Enterococcus villorum/hirae/faecium/faecalis/canintestini/canis/saccharolyticus/ratti/dispar/gallinarum/durans, other Bacteroides, other Gammaproteobacteria}).

@@

@@

@@

>>>>

\title{Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A demonstration of Rmarkdown using Herman Bumpus' data}
\author{Author One\textsuperscript{1},
Author Two\textsuperscript{2},
Author Three\textsuperscript{1,2}}
\date{January 15, 2021}

@@

To be able to better undertand the differences in microbiome profile across all datasets collected, we performed a joint graph-based clustering analysis in order to identify co-regulated bacterial communities (see ``Methods'' section for details). A total of 15 bacterial communities were identified.

Noticebly, bacterial community 15 consisted only of Lactobacillus species (\emph{Lactobacillus coleohominis, Lactobacillus crispatus/acidophilus, Lactobacillus iners, Lactobacillus jensenii, Lactobacillus reuteri/oris/frumenti/antri, other Lactobacillales, Ureaplasma parvum}).

@@

\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-4-1}

@@

\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-6-1}

@@

\includegraphics[width=1\linewidth]{manuscript_template_files/figure-latex/unnamed-chunk-7-1}
